---
title: "Exploring Iris Data Set"
author: "Phil Lombardo"
format: 
  html:
    embed-resources: true
---

## Summary

The iris data set looks at several species of irises, and records some information. We will focus on two variables:

* `Species`: the species of iris
* `Sepal.Width`: the width of the sepal part of the flower.
* `Petal.Width`: the width of the petal part of the flower.

Let's explore.

## Visual
Here is a scatter plot of our main of variables of interest broken up by species.

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 5.5
#| fig-align: center

library(ggplot2)
ggplot(data = iris,
       aes(x = Petal.Width, y = Sepal.Width)) +
  geom_point(aes(color = Species)) + 
  scale_color_brewer(palette = "Set2")+
  scale_y_continuous(breaks = 2:5, limits = c(2, 5.1)) + 
  theme_bw() + 
  labs(x = "Petal Width (in)",
       y = "Sepal Width (in)",
       title = "Comparing Sepal and Petal Widths by Species")
```

The `echo: false` option disables the printing of code (only output is displayed).

### Summary table

```{r}
tapply(iris$Sepal.Width, iris$Species, mean)
```

