---
title: "Exploring Iris and Spotify Data Sets"
author: "Phil Lombardo"
format: 
  html:
    embed-resources: true
    html-table-processing: none
---

```{r}
#| label: setup
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(patchwork)
library(kableExtra)
library(DT)


spotify <- read.csv("data/spotify.csv")

my_table_formatting <- function(table_input, ...){
  table_input %>%
    knitr::kable(...) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover","bordered")) %>%
  column_spec(1, italic = TRUE)
}
```


## Summary

The iris data set looks at several species of irises, and records some information. We will focus on two variables:

* `Species`: the species of iris
* `Sepal.Width`: the width of the sepal part of the flower.
* `Petal.Width`: the width of the petal part of the flower.

Let's explore.

## Iris Visual
Here is a scatter plot of our main of variables of interest broken up by species.

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 5.5
#| fig-align: center

library(ggplot2)
ggplot(data = iris,
       aes(x = Petal.Width, y = Sepal.Width)) +
  geom_point(aes(color = Species)) + 
  scale_color_brewer(palette = "Set2")+
  scale_y_continuous(breaks = 2:5, limits = c(2, 5.1)) + 
  theme_bw() + 
  labs(x = "Petal Width (in)",
       y = "Sepal Width (in)",
       title = "Comparing Sepal and Petal Widths by Species")
```

The `echo: false` option disables the printing of code (only output is displayed).

### Summary table
(Strong recommendation: use `dplyr` to prepare and summarize your data.)

```{r}
#| echo: false

# tapply(iris$Sepal.Width, iris$Species, mean)

iris %>%
  group_by(Species) %>%
  summarise(
    sepal_widths_mean = mean(Sepal.Width),
    sepal_widths_stdev = sd(Sepal.Width),
    sample_size = n()
  ) %>%
  knitr::kable(col.names = c("Species", "Mean Sepal Widths", "Sepal Width Std. Deviations", "Sample Size"),
               align = c('r', 'c', 'c', 'c')) %>%
  kable_styling(full_width = FALSE,
                bootstrap_options = c("hover","bordered")) %>%
  column_spec(1, italic = TRUE) %>%
  # row_spec(1, color='red') %>%
  row_spec(2, background = 'lightblue',bold = TRUE)
```

New summary using `my_table_formatting()`:

```{r}
iris %>%
  group_by(Species) %>%
  summarise(means = mean(Sepal.Length)) %>%
  my_table_formatting(col.names = c("Species", "Mean Sepal Lengths"),
                      align = c('r','c')) %>%
  row_spec(3,color='red',bold = TRUE)
```

### HTML tables with `DT::datatable()` and spotify

```{r}
#| echo: false

spotify %>%
  group_by(key) %>%
  summarise(
    "Mean Streams (millions)" = mean(streams_mill)
    ) %>%
  DT::datatable()
```




# Creating our spotify visual
  
```{r}
#| echo: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7.5

mid_plot <- ggplot(data = spotify,
       aes(x = mode,
           y = danceability,
           fill = mode))+
  geom_boxplot(alpha = .3)+
  geom_jitter(aes(color = mode), alpha = .2, width = .05) +
  scale_fill_manual(values = c("blue","red"))+
  scale_color_manual(values = c("blue","red"))+
  labs(y = "")+
  theme_bw() +
  theme(legend.position = "none")

left_plot <- spotify %>%
  filter(mode == "Major") %>%    # this data frame now only contains "major"
  ggplot(data = .,
         aes(x = danceability)) + 
  geom_histogram(bins = 10,
                 fill = "blue",
                 colour = "white",
                 alpha = .4) +
  labs(x = "danceability",
       y = "Major") +
  scale_y_reverse()+
  coord_flip() +
  theme_bw()

right_plot <- spotify %>%
  filter(mode == "Minor") %>%    # this data frame now only contains "minor"
  ggplot(data = .,
         aes(x = danceability)) + 
  geom_histogram(bins = 10,
                 fill = "red",
                 colour = "white",
                 alpha = .4) +
  labs(x = "",
       y = "Minor") +
  coord_flip() +
  theme_bw()

layout <- "
ABBBC
ABBBC
"
left_plot + mid_plot + right_plot + plot_layout(design = layout) +
  plot_annotation(title = "Dance-ability Score separated by Mode")
```












