---
title: "Exploring Iris Data Set"
author: "Phil Lombardo"
format: 
  html:
    embed-resources: true
---

## Summary

The iris data set looks at several species of irises, and records some information. We will focus on two variables:

* `Species`: the species of iris
* `Sepal.Width`: the width of the sepal part of the flower.
* `Petal.Width`: the width of the petal part of the flower.

Let's explore.

## Visual
Here is a scatter plot of our main of variables of interest broken up by species.

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 5.5
#| fig-align: center

library(ggplot2)
ggplot(data = iris,
       aes(x = Petal.Width, y = Sepal.Width)) +
  geom_point(aes(color = Species)) + 
  scale_color_brewer(palette = "Set2")+
  scale_y_continuous(breaks = 2:5, limits = c(2, 5.1)) + 
  theme_bw() + 
  labs(x = "Petal Width (in)",
       y = "Sepal Width (in)",
       title = "Comparing Sepal and Petal Widths by Species")
```

The `echo: false` option disables the printing of code (only output is displayed).

### Summary table

```{r}
tapply(iris$Sepal.Width, iris$Species, mean)
```

# Creating our spotify visual

```{r}
#| warning: false
#| message: false
#| echo: false

library(tidyverse)
library(patchwork)
spotify <- read.csv("data/spotify.csv")
```
  
```{r}
#| echo: false
#| fig-align: center
#| fig-height: 5
#| fig-width: 7.5

mid_plot <- ggplot(data = spotify,
       aes(x = mode,
           y = danceability,
           fill = mode))+
  geom_boxplot(alpha = .3)+
  geom_jitter(aes(color = mode), alpha = .2, width = .05) +
  scale_fill_manual(values = c("blue","red"))+
  scale_color_manual(values = c("blue","red"))+
  labs(y = "")+
  theme_bw() +
  theme(legend.position = "none")

left_plot <- spotify %>%
  filter(mode == "Major") %>%    # this data frame now only contains "major"
  ggplot(data = .,
         aes(x = danceability)) + 
  geom_histogram(bins = 10,
                 fill = "blue",
                 colour = "white",
                 alpha = .4) +
  labs(x = "danceability",
       y = "Major") +
  scale_y_reverse()+
  coord_flip() +
  theme_bw()

right_plot <- spotify %>%
  filter(mode == "Minor") %>%    # this data frame now only contains "minor"
  ggplot(data = .,
         aes(x = danceability)) + 
  geom_histogram(bins = 10,
                 fill = "red",
                 colour = "white",
                 alpha = .4) +
  labs(x = "",
       y = "Minor") +
  coord_flip() +
  theme_bw()

layout <- "
ABBBC
ABBBC
"
left_plot + mid_plot + right_plot + plot_layout(design = layout) +
  plot_annotation(title = "Dance-ability Score separated by Mode")
```












